% !TeX root = sigui-manual.tex
\newglossaryentry{freshclam}
{
    name={freshclam},
    text={\texttt{freshclam}},
    description={\ClamAV's signature databases updater application}
}
\newglossaryentry{freshclam.conf}
{
    name={freshclam.conf},
    text={\texttt{freshclam.conf}},
    description={The configuration file for \gls{freshclam}}
}
\newglossaryentry{ClamAV}
{
    name={ClamAV},
    text={\textbf{ClamAV}},
    description={Clam AntiVirus engine, see \url{http://www.clamav.net}}
}
\newglossaryentry{db.local.win.clamav.net}
{
    name={db.local.win.clamav.net},
    text={\texttt{db.local.win.clamav.net}},
    description={db.local.win.clamav.net is a round robin record that tries to
	equally balance the traffic between the best database mirrors.}
}
\newglossaryentry{mirror}
{
    name={mirror},
    description={A server holding an exact copy of the original server, for
	better load balancing and bandwidth purposes.}
}
\newglossaryentry{hostname}
{
    name={hostname},
    description={DNS name of a server}
}
\newglossaryentry{CVD}
{
    name={CVD},
    text={\texttt{CVD}},
    description={ClamAV Virus Database. A file that contains multiple signature types, and a digital signature.
	This is the format in which the official signatures are distributed}
}
\newglossaryentry{SigUI}
{
    name={SigUI},
    description={ClamAV for Windows - Signature Configuration User Interface. The
	application documented in this manual}
}
\newglossaryentry{UNC path}
{
    name={UNC path},
    plural={UNC pathes},
    description={Uniform Naming Convention path. A path of the form
	\texttt{\textbackslash\textbackslash ComputerName \textbackslash
	    \textbackslash SharedFolder \textbackslash\textbackslash Resource}, or a long UNC path
	    starting with \texttt{\textbackslash\textbackslash?\textbackslash}}
}
\newglossaryentry{SYSTEM account}
{
    name={SYSTEM account},
    description={A highly privileged account. This is the account used by system
	services. You cannot login as SYSTEM}
}
\newglossaryentry{daily.cvd}
{
    name={daily.cvd},
    description={An important database file for \gls{ClamAV}. Contains often
	updated virus signatures, file type definitions, engine configuration,
	and whitelists}
}
\newglossaryentry{daily.cld}
{
    name={daily.cld},
    description={A \gls{daily.cvd}, after \gls{freshclam} updated it}
}
\newglossaryentry{UAC}
{
    name={UAC},
    description={User Account Control, a security infrastructure introduced in Windows Vista}
}
